### 2016-12-07 paper.tex
Abstract, Intro and comments revised
All around the paper. :D

\documentclass[a4paper,10pt,onecolumn,preprint,3p]{elsarticle}
% The format for draft in Elsevier is one column. ;D

\usepackage{graphics,subfigure}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[dvips]{epsfig}
\usepackage[latin1]{inputenc}
\usepackage{url}

\def\CC{{C\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}}~}
\addtolength{\textfloatsep}{-0.5cm}
\addtolength{\intextsep}{-0.5cm}

\journal{Journal of XXXXXXXXX}

\begin{document}

\begin{frontmatter}

%%%%%%%%%%%%%%%% Title %%%%%%%%%%%%%%%
\title{Predicting Levels of Sales for Newly Published Books \\
using Cost-Sensitive Ensemble Methods}

%%%%%%%%%%%%%%%% Authors %%%%%%%%%%%%%%
\author[abd]{Hossam Faris}
\ead{hossam.faris@ju.edu.jo}
\author[ugrtstc]{Antonio M. Mora}
\ead{amorag@geneura.ugr.es}
\author[ugratc]{Pedro A. Castillo}
\ead{pacv@ugr.es}
\author[ugratc]{J.J. Merelo}
\ead{jmerelo@geneura.ugr.es}
%\author[ugratc]{Pablo Garc\'{\i}a-S\'anchez}
%\ead{pablogarcia@ugr.es}

\address[abd]{Business Information Technology Department, King Abdullah II School for Information Technology \\
The University of Jordan, Amman, Jordan}
\address[ugrtstc]{Department of Signal Theory, Telematics and Communications, ETSIIT and CITIC \\
University of Granada, Granada, Spain}
\address[ugratc]{Department of Computer Architecture and Computer Technology, ETSIIT and CITIC \\
University of Granada, Granada, Spain}

%\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

An important task when publishing a new book is determining an accurate number 
of copies to be printed. Doing both an under- or an overestimation will lead to financial losses.
In this paper, we address the problem of predicting total sales using cost-sensitive ensemble methods.
% Antonio - explain briefly what are them
These can be used as reliable decision-aid tools for publishers, as a part of the final stages of the editorial process. 
Thus, a real dataset including the complete sales data for books published in Spain across several years has been used to compute the prediction models.
% Antonio - extend a bit this with the followed methodology.
Results show that \emph{EXPLAIN THE RESULTS OBTAINED}.

\end{abstract}

\begin{keyword}
% Antonio - to refine later
Book sales forecasting \sep Decision-aid models \sep Feature selection \sep Classification \sep Ensemble methods
\end{keyword}

\end{frontmatter}

%*******************************************************************************
%										INTRODUCTION
%*******************************************************************************
\section{Introduction}
\label{sec:intro}

Inside the publishing process of a new book, deciding the best {\em print run}, i.e. the number of copies to be printed and delivered to bookstores, is a very challenging problem.

To this end, it is extremely important to develop accurate predictive models which could forecast the future sales a new book will achieve, in order to optimise production schedules: to improve the publishing company profits and to minimise losses \cite{Zhao2001}, due both to the excess or to the scarcity of units.
%
%Printing an adequate number of copies, but not too many, might reduce the number 
%of unsold volumes, which could lead to losses. 
%In another hand, if the number of printed copies is not enough, the publisher 
%company might have temporary losses due to a lack of supply.
%
Since, according to the results reported in \cite{Fildes2010,Saeed2008}, reducing the error in estimated sales leads to an enhancement of the whole production process.

Usually the prediction of book sales has been generally based on experts' experience. They analyse data about sales and, taking into account their knowledge of the industry, the production process, and their perception about trends, and take decisions on the production and publication process.
However, this process might be tedious, especially when the amount of data is 
quite large. Moreover, the results are not robust, since different experts can reach different predictions from the same dataset \cite{Sanders1994}.

\textbf{UNDER CONSTRUCTION}
% Antonio - I remove the ';)' because I want to check the parentheses correspondence and it breaks it. :P

To make the print run as close as possible to the actual level of sales, it is 
essential to have an estimation of those sales. 
In this sense, this paper continues our previous work on book sales prediction 
\cite{Castillo2016books} in which the main factors influencing sales where 
analyzed in order to create a tool that the publisher can use to decide how many 
books should be printed.
In that paper, several regression methods were applied to estimate sales for new launched books using a real dataset of book sales related to the Spanish market, provided by the Spanish publishing company Trevenque Editorial S.L.\footnote{\tt http://editorial.trevenque.es}

However, as finding adequate forecasting techniques and selecting the best method 
to use represent also a problem, we propose configuring complex models based on
ensembles \cite{Qiu2014} in order to face the challenge of determining the actual 
level of sales a new book will achieve.
Moreover, since publishers need models to explain obtained predictions, and 
`black box' forecasting methods may not be suitable, using these ensemble 
models publishers will be able to combine their expert knowledge about the 
market with the created forecasting models in order to get a reliable estimation 
of book sales level.

% Antonio - We must remark that this approach is based in classification rather than regression, which is based on levels of sales (classes) instead of on numerical results. These will be easily 'explained' or justified for a (human) publisher.
% Previous work also presented the results in an intelligible way, since it offered decision trees as the last output, but they include regression computation in the leaves. Here the models/outputs will be easily interpreted.

The rest of this paper is structured as follows: 
A detailed review of the approaches related to sales prediction in similar scopes is presented in Section \ref{sec:soa}.
Then, Section \ref{sec:methodology} details the methodology considered in the present study. 
Obtained results after extensive experimentation are presented in Section \ref{sec:experiments_results}.
Finally, conclusions and future lines of work are presented in Section \ref{sec:conclusions_future_work}.

%*******************************************************************************
%										STATE OF THE ART
%*******************************************************************************
\section{State of the art on sales forecasting} 
\label{sec:soa}

\textbf{UNDER CONSTRUCTION}
% Antonio - I remove the ';)' because I want to check the parentheses correspondence and it breaks it. :P

%*******************************************************************************
%										METHODOLOGY
%*******************************************************************************
\section{Methodology}
\label{sec:methodology}

\begin{figure*}[ht]
\begin{center}
\includegraphics[scale=0.40]{books_methodology}
\end{center}
\caption{Proposed methodology.}
\label{fig:methodology}
\end{figure*}
% Antonio - I like this figure, however, it seems on it that the results of the basic classifier are used as 'starting point' of the ensembles, and the results of these methods are used in the cost sensitive classification... is this correct? Otherwise, maybe we should change the design of the figure. ;D
% Hossam: I have changed the figure as Antonio suggested ;)
% Antonio - It is better now, however, is the evaluation the last stage of the methodology? Won't we choose the best approach based on the results? Won't we compose an useful tool with it as in previous paper? I mean a decision tree or so. ;D

%*******************************************************************************
%										MATERIALS AND METHOD
%*******************************************************************************
\section{Materials and methods}
\label{sec:materials}

%-------------------------- BASIC CLASSIFIERS --------------------------
\subsection{Basic classifiers}
\label{subsec:classifiers}

In order to show the need for more intelligent and complex classification models, basic and more common models are experimented and applied first. 
\begin{itemize}

\item \textbf{$k$-Nearest Neighbors ($k$-NN)}
The kNN \cite{Aha1991,Mitchell1997} method is an instance-based classifier in 
which unknown patterns/instances are classified by relating them to the labelled ones, using a distance measure or function, usually Euclidean or Minkowsky distances.

The algorithm is based on the idea that two patterns with a high distance between them are less likely to belong to the same class than two close ones.

In the simplest approach, for every unknown instance to classify, the algorithm finds its nearest neighbour (in the patterns space), and assigns the class of that neighbour to the unknown pattern.
However, it is more common to apply a more robust approach, in which $k$ neighbours are located, and thus, assigning the majority class of these instances to the unclassified pattern.

\item \textbf{Naive Bayes classifier (NB)}

This is a type of probabilistic classifier, based on the Bayes' Theorem \cite{AI-book_Bayes}, which considers a \textit{naive} (or strong) independence between the variables of the samples to be classified.

It is normally used as a baseline for comparison, because it usually yields competitive results in low computational time. Moreover, naive Bayes classifier shows a very good performance when there is considered a supervised training stage.

This method has a very good scalability, requiring a linear number of parameters to work, depending on the number of variables of the problem. These parameters are the mean and standard deviations of the variables of every class.
Thus, it just evaluates closed-form expressions, not needing to iterate to refine the solutions as most of methods do.

\item \textbf{J48 - Decision Trees (DT)}

This method is an implementation of the C4.5 classifier, which generates the so-called \textit{decision trees}. It was introduced in 1993 by Quinlan \cite{Quinlan1993} as an extension of the ID3 algorithm \cite{Mitchell1997}.

The algorithm builds a decision tree using the attributes of the input patterns as nodes and considering its values for defining a decision in that node. Thus, for each node, the best attribute for splitting the training set is selected. The criteria for choosing one attribute is based on the difference of entropy (normalized information gain) obtained by every possible division. The one which yields the highest information gain is chosen as the best. This process is repeated for the subsequent data subsets, building subtrees. Normally, the decision tree is iteratively refined (or pruned) to maximize these gains and to avoid useless subtrees.

\item \textbf{Multilayer Perceptron Neural Network (MLP)}

It is an artificial neural network generally used for classification or 
approximation/regression problems \cite{Rosenblatt1962,Widrow1990}. It maps the input data onto an appropriate output, as a generalisation of the standard linear perceptron that uses several layers of nodes, called neurons. Each neuron consists of a linear combination of weighted inputs which is passed through a non-linear activation function to produce its output.

Thus, a MLP is able to solve linearly inseparable problems \cite{SteinwenderBitzer2003}. It is normally trained using a supervised learning 
technique called \textit{back-propagation}. This method was proposed by Werbos, Parker and Rumelhart et al. \cite{Werbos1974,Parker1985,Rumelhart1985}, and consists in updating the weights of the output layer neurons according to the obtained erroneous output. Then, the weights are updated inversely up to the input layer.
There are normally two hidden layers, as some works \cite{Lippmann1987,Bishop1996} demonstrated to be enough to 
create classifying regions of any kind.

\item \textbf{Support Vector Machines (SVM)}
SVM \cite{Cortes1995,Shevade2000} is a method which applies statistical learning theory. In classification problems, the algorithm searches for an optimal hyperplane that separates two classes, maximising the margin between them. 
This hyperplane is defined by a subset of training set samples, called support vectors.

This method is highly robust and reliable even if the space is highly dimensional and the problem is not linearly separable.

SVM has been successfully applied in forecasting and classification problems \cite{Cao2003,MinSVM05,Jari2008}.

\end{itemize}

%-------------------------- ENSEMBLE CLASSIFIERS --------------------------
\subsection{Ensemble classifiers}
\label{subsec:ensembles}

\begin{itemize}
\item \textbf{Bagging}: 
It is also called \textit{Bootstrap aggregating} is an ensemble meta-algorithm introduced in \cite{B1996}. Bagging works by training base classifiers (usually decision tree classifiers) based on generated new datasets from the original training dataset. The new generated datasets are drawn randomly with replacement in a size almost equal to the original dataset. 
% Antonio - I cannot understand this sentence (upper one). :/
The final classifier's output is based on aggregating the base classifiers by taking a majority or weighted vote.

\item \textbf{Boosting}: 
It is an ensemble classification algorithm which was firstly proposed in \cite{schapire1990strength}. Boosting is founded on the idea that a set of weak classifiers can create a stronger one. In this work we adopt a powerful a variation of boosting called \textit{Adaptive Boosting} or ``AdaBoost'' for short \cite{FS1997}. AdaBoost operates on a training dataset ${(x_{1},y_{1}),...,(x_{m},y_{m})}$ where each $x_{i}$ is a set of inputs belonging to some domain $X$ and each $y_{i}$ belongs to some label $Y$ assuming $Y$ is a binary class set $\{+1,-1\}$.

During the learning process, AdaBoost modifies the dataset by weighting its samples. These weights are updated every time a new classifier is trained to increase the focus 
% Antonio - what is the focus here?
on the difficult data instances. Consequently, the final generated classifier is a weighted combination of all trained base classifiers. According to many previous works, 
% Antonio - if you say this, you must add some references. ;P
AdaBoost is not affected 
% Antonio - 'is not affected by' -> 'avoids'?
by overfitting and it can outperform many other state-of-the-art classification techniques.

\item \textbf{Random Forest}:
This method is based on the construction of a set of decision trees using a stochastic process over the basis of C4.5 algorithm. It was proposed by Breiman in \cite{Breiman2001} and aims to create independent and uncorrelated trees based on different and random input vectors, following the same distribution. 
The result is the average of the generated trees during the process.

% Antonio - we should improve a bit this description to do it more complete. ;)

\end{itemize}

%*******************************************************************************
%										COST SENSITIVE CLASSIFICATION
%*******************************************************************************
\section{Cost sensitive classification}
\label{sec:cost_sensitive_classification}

%------------------------ COST SENSITIVE LEARNING ------------------------
\subsection{Cost sensitive learning}
\label{subsec:cost_learning}

%------------------------ COST SENSITIVE CLASSIFICATION ------------------------
\subsection{Cost sensitive classification}
\label{subsec:cost_classification}

%*******************************************************************************
%										DATASET
%*******************************************************************************
\section{Dataset description}
\label{sec:dataset}

The obtained dataset contains collected information about 6083 different books produced by 209 publishers. Each book or pattern in the dataset is represented by 10 features as shown in Table \ref{tabla:params_pre_sales}.

\begin{table*}
\caption{Pre-sales features and Sales level(Output class). } 
\label{tabla:params_pre_sales}
\begin{center}
\begin{tabular}{|c|l|l|c|c|}
\hline 
No. & Feature Name & Variable & Type & In/Out\\
\hline 
1 & retail price (when launched) & \texttt{ret\_price} & numerical & input\\
2 & main subject (code) & \texttt{subject1} & numerical & input\\
3 & bookbinding (code) & \texttt{binding} & categorical & input\\
4 & gifts (promotional books) & \texttt{gifts} & numerical & input\\
5 & units distributed as novelty & \texttt{distrib\_novelty} & numerical & input\\
6 & total number of points of sale & \texttt{tot\_points\_sale} & numerical & input\\
7 & total number of points of sale (1st year) & \texttt{tot\_points\_sale\_1st\_year} & numerical & input\\
8 & weeks on sale & \texttt{weeks\_sale} & numerical & input\\
9 & print run & \texttt{print\_run} & numerical & input\\
\hline 
\hline
10 & Sales level & \texttt{sales\_level} & categorical & output\\
\hline 
\end{tabular}
\end{center}
\end{table*}

The output variable which is related to the total amount of sales is a categorical variable which represent the level of the quantity sold for a given book. Thus, four categories have been considered to represent these levels: \textit{Low (L)}, \textit{Medium (M)}, \textit{High (H)}, and \textit{Very High (VH)}. Table \ref{table:freq} shows the ranges of total sales that correspond to each category along with the number of sold books that fall in. The discretization of the volume of sales was conducted by an expert in the domain of editorial management business, who defined the four categories.

Table \ref{table:freq} reveal more information about the nature of the dataset. We can notice that most of the investigated books are categorized as Low and Medium levels forming 26.8\% and 36.4\% of the dataset, respectively. High volumes of sold books that range from 300 up to 1000 are categorized as High volume of sales and form 24\% of the dataset. On the other hand, books that achieved more than 1000 sold copies are categorized as Very High and they form around 12.8\% of the dataset. Our focus in this study is to identify the more profitable categories which are High and Very High. However, since the latter category is much smaller than the other ones, we can describe our dataset as imbalanced dataset which makes the problem of identifying this category harder and more challenging.

\begin{table*}[ht]
\caption{Categories corresponding to total sales }
\centering{}%
\begin{tabular}{|c|c|c|c|}
\hline 
Category & Range of total sales & Count & Ratio\tabularnewline
\hline 
\hline 
Low & 0-60 & 1631 & 26.8\%\tabularnewline
\hline 
Medium & 61-300 & 2212 & 36.4\%\tabularnewline
\hline 
High & 300-1000 & 1460 & 24\%\tabularnewline
\hline 
Very High & $>$1000 & 780 & 12.8\%\tabularnewline
\hline 
\end{tabular}
\label{table:freq}
\end{table*}

%*******************************************************************************
%										EVALUATION MEASUREMENTS
%*******************************************************************************
\section{Evaluation measurements}
\label{sec:eval_measures}

To evaluate the developed classification models, first, we refer the confusion matrix shown in Table \ref{fig:confusionmatrix} which is considered as a basic source for evaluating any binary classification model. In our case which is a multiclass prediction problem, this matrix can be expanded with a row and column for each class to evaluate the multiclass classification models. Three measurements are used to evaluate the applied classifiers: accuracy rate, precision and recall. The accuracy rate simply measures the total number of correctly classified instances over the total number of instances $N$ as shown in Equation \ref{equation:acc}.

Precision is the ratio of the correctly classified instances as $i$ to the number of instances classified as $i$. On the other side, recall is the ratio of the correctly classified instances as $i$ to the number of actual $i$ instances. Precision and recall are given in Equation \ref{equation:precision} and \ref{equation:recall}, respectively.

\begin{figure*}[ht]
\begin{center}
\includegraphics[scale=0.50]{Confusion-matrix}
\end{center}
\caption{Basic confusion matrix.}
\label{fig:confusionmatrix}
\end{figure*}

\begin{equation}
Accuracy=\frac{\sum_{i}C_{ii}}{N}
\label{equation:acc}
\end{equation}

\begin{equation}
Precision_{i}=\frac{C_{ii}}{\sum_{j}C_{ij}}
\label{equation:precision}
\end{equation}

\begin{equation}
Recall_{i}=\frac{C_{ii}}{\sum_{j}C_{ji}}
\label{equation:recall}
\end{equation}

\begin{figure*}[ht]
\begin{center}
\includegraphics[scale=0.40]{Confusion-matrix-multiclass-default}
\end{center}
\caption{Default cost matrix.}
\label{fig:Default_cost_matrix}
\end{figure*}

%*******************************************************************************
%										EXPERIMENTS AND RESULTS
%*******************************************************************************
\section{Experiments and results}
\label{sec:experiments_results}

%This is just the first step of the experiments
% This table shows the average results of 10X10-folds experiments of the basic classifiers.
% It is impossible to calculate directly the average Precision and Recall for each class for multiclass problems in Weka Experimenter. So I have written a bash script to generate the row predictions utilizing Weka 3.9 developers version. Then I wrote another Python script to process these predicitons and calculate the precision and recall measurements as listed in this table.

In all experiments, 10-folds cross validation methodology is applied for training and testing the developed classifiers. Moreover, in order to generate statistically reliable results, the 10-fold cross validation is repeated 10 times with different partitioning seeds then the average and standard deviation are calculated for each evaluation metric.

In the first stage of the experiments the basic classifiers k-NN, NB, MLP, J48 and SVM are applied for predicting the sales level. The performance of some classifiers is highly affected by the initial values of their parameters. For this reason, initial experiment were conducted to tune the parameters of k-NN, SVM and MLP. For k-NN the best classification results were obtained with k=1. For SVM, a rough grid search was used, the value for the cost $C$ is set to 50 and $\gamma$  to 0.01. For MLP, the number of hidden neurons is set to (number of features+ number of output classes)/2.

The developed classifiers are assessed by measuring their accuracy and the precision and recall of each class label. Evaluation measures for all classifiers are shown in Table \ref{table:basic_classifiers}.

\begin{table*}
\caption{Classification results using basic classifiers (without print-run)}
\centering{}%
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
 & kNN  & NB  & J48  & MLP  & SVM\tabularnewline
\hline 
\hline 
$Accuracy$  & 0.69$\pm$0.0026 & 0.67$\pm$0.0008 & 0.7$\pm$0.004 & 0.71$\pm$0.0027 & 0.67$\pm$0.0012\tabularnewline
\hline 
\hline 
$Recall_{L}$  & 0.75$\pm$0.0044 & 0.86$\pm$0.0013 & 0.79$\pm$0.0071 & 0.78$\pm$0.0222 & 0.59$\pm$0.0035\tabularnewline
\hline 
$Precision_{L}$  & 0.78$\pm$0.003 & 0.73$\pm$0.0017 & 0.8$\pm$0.0049 & 0.8$\pm$0.0173 & 0.84$\pm$0.0025\tabularnewline
\hline 
\hline 
$Recall_{M}$  & 0.74$\pm$0.0034 & 0.66$\pm$0.0024 & 0.7$\pm$0.0072 & 0.71$\pm$0.0129 & 0.82$\pm$0.002\tabularnewline
\hline 
$Precision_{M}$  & 0.65$\pm$0.003 & 0.64$\pm$0.0007 & 0.68$\pm$0.0079 & 0.69$\pm$0.007 & 0.59$\pm$0.0012\tabularnewline
\hline 
\hline 
$Recall_{H}$  & 0.59$\pm$0.0047 & 0.53$\pm$0.0022 & 0.6$\pm$0.0099 & 0.65$\pm$0.0346 & 0.55$\pm$0.0016\tabularnewline
\hline 
$Precision_{H}$  & 0.62$\pm$0.0037 & 0.58$\pm$0.0017 & 0.62$\pm$0.0058 & 0.62$\pm$0.0109 & 0.63$\pm$0.0012\tabularnewline
\hline 
\hline 
$Recall_{VH}$  & 0.6$\pm$0.0073 & 0.53$\pm$0.003 & 0.72$\pm$0.0086 & 0.68$\pm$0.0336 & 0.61$\pm$0.0012\tabularnewline
\hline 
$Precision_{VH}$  & 0.79$\pm$0.0071 & 0.81$\pm$0.0032 & 0.74$\pm$0.0072 & 0.77$\pm$0.0212 & 0.78$\pm$0.0028\tabularnewline
\hline 
\end{tabular}
\label{table:basic_classifiers}
\end{table*}

It can be noticed that all classifiers have obtained very close accuracy ratios with a slight advantage for the MLP network with 71\% accuracy. In general, accuracy rates don't give an insight on how the classifier performs regarding each class. Therefore, precision and recall rates are examined for each class. Precision and recall values in Table \ref{table:basic_classifiers} show that J48 classifier achieved the highest recall value of 72\% for (VH) class with 74\% precision. The second best classifier for this important class is MLP with 68\% recall and 77\% precision. Having a look at the second important class which is (H), we can see that MLP has the highest recall with 65\% and a precision of 62\%. J48 comes second for the (H) class with a slight decrease in the recall value.

In the next stage of the experiments, the ensemble classifiers (i.e. RF, Bagging and AdaBoost) are applied and evaluated. For Bagging and AdaBoost, the base classifier is selected to be the J48 decision tree algorithm. This selection is for three reasons: first, the J48 algorithm achieved very competitive evaluation results in the previous phase. Second, J48 is much faster as a learning algorithm compared to the MLP network which was the second candidate. Third, tree induction algorithms are good candidates as base classifiers in bagging and boosting because support diversity among the learners. It is well known that tree induction algorithms are relatively unstable (i.e. weak) base learners, meaning that small changes in the training data lead to big changes in the developed models.

\begin{figure*}

\centerline{\includegraphics[scale=0.46]{RF}}

\caption{Performance of RF classifier over different number of trees.}
\label{fig:PerformanceRF}
\end{figure*}

\begin{figure*}
\centerline{
\includegraphics[scale=0.46]{Bagging}}

\caption{Performance of Bagging classifier over different number of iterations.}
\label{fig:PerformanceBagging}
\end{figure*}

\begin{figure*}
\centerline{
\includegraphics[scale=0.46]{Adaboost}
}
\caption{Performance of Adaboost classifier over different number of iterations.}
\label{fig:PerformanceAdaboost}
\end{figure*}

\begin{table*}
\caption{Classification results using ensemble classifiers (RF, AdaBoost and Bagging)}
\centering{}%
\begin{tabular}{|c|c|c|c|}
\hline 
 & RF  &  AdaBoost &  Bagging \tabularnewline
\hline 
 \hline $Accuracy$ & 0.741 $\pm$ 0.002 & 0.743 $\pm$ 0.0023 & 0.735 $\pm$ 0.0022 \tabularnewline 
 \hline 
 \hline $Recall_{L}$ & 0.796 $\pm$ 0.0038 & 0.793 $\pm$ 0.0038 & 0.792 $\pm$ 0.0048 \tabularnewline 
 \hline $Precision_{L}$ & 0.823 $\pm$ 0.0035 & 0.827 $\pm$ 0.0032 & 0.814 $\pm$ 0.0043 \tabularnewline 
 \hline 
 \hline $Recall_{M}$ & 0.748 $\pm$ 0.0033 & 0.75 $\pm$ 0.0029 & 0.738 $\pm$ 0.0038 \tabularnewline 
 \hline $Precision_{M}$ & 0.712 $\pm$ 0.0026 & 0.716 $\pm$ 0.0032 & 0.707 $\pm$ 0.004 \tabularnewline 
 \hline 
 \hline $Recall_{H}$ & 0.68 $\pm$ 0.0073 & 0.684 $\pm$ 0.0051 & 0.667 $\pm$ 0.0065 \tabularnewline 
 \hline $Precision_{H}$ & 0.666 $\pm$ 0.0037 & 0.667 $\pm$ 0.0029 & 0.663 $\pm$ 0.004 \tabularnewline 
 \hline 
 \hline $Recall_{VH}$ & 0.723 $\pm$ 0.0054 & 0.727 $\pm$ 0.0041 & 0.731 $\pm$ 0.0057 \tabularnewline 
 \hline $Precision_{VH}$ & 0.816 $\pm$ 0.0051 & 0.808 $\pm$ 0.0056 & 0.793 $\pm$ 0.0069 \tabularnewline 
\hline 
\end{tabular}
\label{table:ensemble_classifiers}
\end{table*}

%*******************************************************************************
%										CONCLUSIONS AND FUTURE WORK
%*******************************************************************************
\section{Conclusions and Future Work}
\label{sec:conclusions_future_work}

%********************************************************************************
\section*{Acknowledgements}

This work has been supported in part by projects PreTEL (PRM Consultores - Trevenque S.L.), TIN2014-56494-C4-3-P and TEC2015-68752 (Spanish Ministry of Economy and Competitiveness and FEDER), and PROY-PP2015-06 (Plan Propio 2015, funded by the University of Granada, Spain).

%********************************************************************************
\bibliographystyle{elsarticle-num}
\bibliography{refs}

%----------------------------------------------------------------------

\end{document}